---
title: "423_final"
output: pdf_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
library(MASS)
library(ggplot2)
library(corrplot)
library(GGally)
```

```{r}
# Data Preparation
sleep_df <- read_csv('/Users/edisonlu/Desktop/STAT 423/sleep_cycle_productivity.csv',
                     show_col_types = FALSE)

sleep_df <- sleep_df %>%
  mutate(
    `Sleep Start Time` = hms::hms(round(as.numeric(`Sleep Start Time`) * 3600)),
    `Sleep End Time` = hms::hms(round(as.numeric(`Sleep End Time`) * 3600)),
    Age = as.factor(Age),
    Gender = as.factor(Gender),
    `Sleep Quality` = as.factor(`Sleep Quality`),
    `Productivity Score` = as.factor(`Productivity Score`),
    `Mood Score` = as.factor(`Mood Score`),
    `Stress Level` = as.factor(`Stress Level`)
  )

summary(sleep_df)
```

```{r}
# Elementary Data Analysis
hist(sleep_df$`Total Sleep Hours`,
     main = "Distribution of total sleep hours",
     xlab = "Total Sleep Hours",
     col = "purple", border = "gold")

# Plot distribution of interested parameters
par(mfrow = c(2,2))

plot_fun <- function(var, title, xtitle) {
  plot <- hist(var, 
     main = title,
     xlab = xtitle,
     col = "lightblue", border = "black")
  
  return(plot)
}

plot_fun(sleep_df$`Exercise (mins/day)`, "Histogram of Exercise Minutes", "Exercise (mins/day)")
plot_fun(sleep_df$`Caffeine Intake (mg)`, "Histogram of Caffeine Intake (mg)", "Caffeine Intake (mg)")
plot_fun(sleep_df$`Screen Time Before Bed (mins)`, "Histogram of sleep time before bed (mins)", "Screen time before time (mins)")
plot_fun(sleep_df$`Work Hours (hrs/day)`, "Histogram of Work hours", "Work hours")
```
```{r}
# Some more EDA
sleep_df_grouped <- sleep_df %>%
  group_by(Age) %>%
  summarise(mean_sleep_hours = mean(`Total Sleep Hours`), .groups = "drop")

ggplot(data = sleep_df_grouped,
       mapping = aes(x = Age, y = mean_sleep_hours)) +
  geom_point(size = 2) +
  labs(
    title = "Mean total sleep hours by age",
    x = "Age",
    y = "Mean sleep hours"
  ) +
  theme_minimal()
```


```{r}
# Explore relationship between the response variable and predicting variables we are interested in
plot(x = sleep_df$`Exercise (mins/day)`,
     y = sleep_df$`Total Sleep Hours`,
     main = "Scatterplot of exercise and total sleep hours",
     xlab = "Exercise (mins/day)",
     ylab = "Total Sleep Hours")

plot(x = sleep_df$`Caffeine Intake (mg)`,
     y = sleep_df$`Total Sleep Hours`,
     main = "Scatterplot of Caffeine Intake and total sleep hours",
     xlab = "Caffeine Intake",
     ylab = "Total Sleep Hours")

plot(x = sleep_df$`Screen Time Before Bed (mins)`,
     y = sleep_df$`Total Sleep Hours`,
     main = "Scatterplot of Screen Time Before Bed and total sleep hours",
     xlab = "Screen Time Before Bed",
     ylab = "Total Sleep Hours")
```

After completing some preliminary data exploration, it has come to my attention that the main four interested variables, namely exercise per day (unit: minutes), caffeine intake (unit: mg), screen time before bed time (unit: minutes), and work hours does not show any sign of skewness in their respective distributions, hence it is unnecessary to conduct transformation on them.

```{r}
# Construct the full multiple linear regression model
full_model <- lm(`Total Sleep Hours` ~ Age + Gender + `Sleep Quality` + `Exercise (mins/day)` + `Caffeine Intake (mg)` + `Screen Time Before Bed (mins)` + `Work Hours (hrs/day)` + `Productivity Score` + `Mood Score` + `Stress Level`, data = sleep_df)

summary(full_model)
```

```{r}
# Apply Box-Cox transformation
boxcox_result <- boxcox(full_model, lambda = seq(-2, 2, by = 0.1))
boxcox_result

optimal_lambda <- boxcox_result$x[which.max(boxcox_result$y)]
cat("The optimal lambda value is", optimal_lambda)

abline(v = optimal_lambda, col = "red", lwd = 2)
text(optimal_lambda, par("usr")[4] - 50, labels = round(optimal_lambda, 3), col = "red", pos = 4)
```
```{r}
# Transform the response variable
# sleep_df$`Total Sleep Hours` <- sapply()
```


```{r}
# ANOVA on whether gender has influence on sleep quality
lm.fit_gender <- lm(`Total Sleep Hours` ~ Gender, data = sleep_df)
anova_gender <- anova(lm.fit_gender, full_model)

print(anova_gender)

# ANOVA on whether age has influence on sleep quality
lm.fit_age <- lm(`Total Sleep Hours` ~ Age, data = sleep_df)
anova_age <- anova(lm.fit_age, full_model)

print(anova_age)
```

```{r}
# Feature selection using ANOVA F-test
step_model <- step(full_model, direction = "both")
summary(step_model)
```

```{r}
# ANOVA F-test
anova(step_model, full_model)
```

```{r}
# Check for model assumptions
par(mfrow = c(2,2))

plot(full_model)

cooksd <- cooks.distance(full_model)
plot(cooksd, type = "h", col = "black", lwd = 2, ylab = "Cook's distance", 
     xlab = "Observation number", main = "Cook’s Distance Plot")
abline(h = 4/length(cooksd), col = "red", lty = 2)  # Add threshold line
```

```{r}
res <- residuals(full_model)
hist(res, breaks = 30,
     prob = TRUE,
     main = "Histogram of residuals of full model",
     xlab = "Residuals",
     ylab = "Density",
     col = "purple", border = "gold")

curve(dnorm(x, mean = mean(res), sd = sd(res)), col = "black", lwd = 2, add = TRUE)
```

```{r}
# Check for multicollinarity
cor_matrix <- sleep_df %>%
  dplyr::select(`Total Sleep Hours`, `Exercise (mins/day)`, `Caffeine Intake (mg)`,
         `Screen Time Before Bed (mins)`, `Work Hours (hrs/day)`) %>%
  cor(use = "complete.obs")

corrplot(cor_matrix, method = "color", type = "upper",
         col = colorRampPalette(c("blue", "white", "red"))(200),
         tl.cex = 0.8, tl.col = "black", addCoef.col = "black", number.cex = 0.7)
```