---
title: "STAT423 Final Project"
author: "Chenyi Wang"
date: "2025-03-07"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# load library
```{r}
library(ggplot2)
library(dplyr)
library(car)
library(MASS)
```

# load the data
```{r}
sleep_data <- read.csv("/Users/macbook/Downloads/sleep_cycle_productivity.csv")
```

# check for missing values
```{r}
colSums(is.na(sleep_data))
```
# Ensure all variables are numeric and within reasonable ranges
```{r}
summary(sleep_data)
```
# Standardize data
```{r}
sleep_data <- sleep_data %>%
  mutate(
    Stress_Level = scale(Stress.Level),
    Sleep_Quality = scale(Sleep.Quality),
    Caffeine_Intake = scale(Caffeine.Intake..mg.)
  )
```

# Scatter plot for stress level vs. sleep quality
```{r}
ggplot(sleep_data, aes(x = Stress_Level, y = Sleep_Quality)) +
  geom_jitter(width = 0.2, height = 0.2, alpha = 0.5) +
  labs(title = "Stress Level vs. Sleep Quality", 
       x = "Stress Level", 
       y = "Sleep Quality")
```

# Scatter plot for caffeine intake vs. sleep quality
```{r}
ggplot(sleep_data, aes(x = Caffeine_Intake, y = Sleep_Quality)) +
  geom_jitter(width = 0.2, height = 0.2, alpha = 0.5) +
  labs(title = "Caffeine Intake vs. Sleep Quality",
       x = "Caffeine Intake (mg)",
       y = "Sleep Quality")
```
# Summary Statistics
```{r}
summary_stats <- sleep_data %>%
  summarise(
    Stress_Mean = mean(Stress_Level),
    Stress_Median = median(Stress_Level),
    Stress_SD = sd(Stress_Level),
    Sleep_Mean = mean(Sleep_Quality),
    Sleep_Median = median(Sleep_Quality),
    Sleep_SD = sd(Sleep_Quality),
    Caffeine_Mean = mean(Caffeine_Intake),
    Caffeine_Median = median(Caffeine_Intake),
    Caffeine_SD = sd(Caffeine_Intake)
  )
print(summary_stats)
```
# Stratify by caffeine intake levels (low, moderate, high)
```{r}
sleep_data <- sleep_data %>%
  mutate(Caffeine_Level = case_when(
    Caffeine_Intake < -0.5 ~ "Low",
    Caffeine_Intake >= -0.5 & Caffeine_Intake <= 0.5 ~ "Moderate",
    Caffeine_Intake > 0.5 ~ "High"
  ))
```
# Compare means of stress level and sleep quality between caffeine groups
```{r}
caffeine_groups <- sleep_data %>%
  group_by(Caffeine_Level) %>%
  summarise(
    Mean_Stress = mean(Stress_Level),
    Mean_Sleep = mean(Sleep_Quality)
  )
print(caffeine_groups)
```

# Hypothesisi Testing
```{r}
model <- lm(Sleep_Quality ~ Stress_Level * Caffeine_Intake, data = sleep_data)
summary(model)
```
# Residuals vs. Fitted values plot
```{r}
ggplot(data = data.frame(fitted = fitted(model), residuals = residuals(model)), 
       aes(x = fitted, y = residuals)) +
  geom_point(alpha = 0.3) +
  geom_hline(yintercept = 0, color = "red") +
  labs(title = "Residuals vs. Fitted Values", x = "Fitted Values", y = "Residuals") 
```
# QQ plot of residuals
```{r}
qqnorm(residuals(model))
qqline(residuals(model), col = "red")
```
# Check for multicollinearity using VIF
```{r}
vif_values <- vif(model, type = "predictor")
print(vif_values)
```


